% !TEX program = LuaLaTeX+se
\documentclass[12pt,a4paper]{article}
\usepackage[titletoc,toc,title]{appendix}
\usepackage{fontspec}
\defaultfontfeatures{Ligatures=TeX}
\setmainfont[
    BoldFont=LinLibertineOZ,
    BoldItalicFont=LinLibertineOZI,
    SmallCapsFont=LinLibertineO,
    SmallCapsFeatures={Letters=SmallCaps},
]{LinuxLibertineO}
\setsansfont[Extension=.otf,
    BoldFont=LinBiolinum_RB,
    ItalicFont=LinBiolinum_RI,
    BoldItalicFont=LinBiolinum_RB,% fake
    SmallCapsFont=LinBiolinum_R,
    SmallCapsFeatures={Letters=SmallCaps},
    ]{LinBiolinum_R}
\setmonofont[Extension=.otf,
    ItalicFont=Inconsolatazi4-Regular,
    BoldFont=Inconsolatazi4-Bold,
    BoldItalicFont=Inconsolatazi4-Bold,
    AutoFakeSlant,
    ItalicFeatures={FakeSlant},
    BoldItalicFeatures={FakeSlant}
    ]{Inconsolatazi4-Regular}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage{multicol}
\usepackage{units}
\usepackage{framed}
\usepackage{url}
\usepackage{tabulary}
\usepackage{tabularx}

\usepackage{makeidx}
\makeindex

\usepackage[table]{xcolor}
\definecolor{lightgray}{gray}{0.9}
\definecolor{green}{HTML}{0c700c}
\definecolor{myred}{HTML}{FF3333}

\usepackage{minted} % must be after xcolor
\newminted{latex}{gobble=1,bgcolor=lightgray}% Originally had 'gobble=2' but this creates problems on systems where the 2-space indentation is switched out for single tabs (Br. Samuel's, for one, does this).  Should be set to autogobble once every developer is using TeXLive 2015 or later (as the autogobble feature doesn't appear in TeXLive 2014 or earlier).
%\newminted{gabc}{bgcolor=lightgray} % can this be done?
\newenvironment{gabccode}{\tt}{}


\usepackage{gregoriotex}

\usepackage{carolmin}
\usepackage{mflogo}

\usepackage{hyperref}
\hypersetup{colorlinks, citecolor=black, filecolor=black, linkcolor=green, urlcolor=green}

\usepackage{longtable}
\usepackage{multirow}
\usepackage{pdflscape}
\usepackage{hhline}
\usepackage{listings}
\directlua{dofile('GregorioRef.lua')}

\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\nb}{n.b.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\newcommand*{\etc}{etc.\@\xspace}

\newcommand{\macroname}[3]{%
  \vspace{3.25ex plus 1ex minus .2ex}%
  \pagebreak[3]%
  \makebox[\linewidth]{\ttfamily\bfseries #1#2%
  \hspace{\fill}\normalfont\itshape #3}%
  \vspace{1.5ex plus .2ex}%
  \index{#1}}
  %\addcontentsline{toc}{subsubsection}{#1}}

\newcommand{\optional}[1]{{\itshape #1}}

\lstset{backgroundcolor=\color{lightgray},
        basicstyle=\small\ttfamily,
        numbers=left,
        numberstyle=\footnotesize,
        stepnumber=1,
        numbersep=5pt}

% for the character tables
\font\greciliae = {name:greciliae} at 1000000 sp\relax
\font\gregorio = {name:gregorio} at 1000000 sp\relax
\font\parmesan = {name:parmesan} at 1000000 sp\relax
\font\greciliaeOp = {name:greciliae-op} at 1000000 sp\relax
\font\gregorioOp = {name:gregorio-op} at 1000000 sp\relax
\font\parmesanOp = {name:parmesan-op} at 1000000 sp\relax
\font\greextra = {name:greextra} at 12 pt\relax
\newcommand{\excluded}[1]{{\tiny\itshape #1}}

\newenvironment{argtable}{%
  \bigskip\rowcolors{1}{lightgray}{lightgray}
  \tabularx{\textwidth}{clX}
    Arg & Value & Description \\
    \hline}%
  {\endtabularx\bigskip}

\setlength{\parindent}{0mm} % Default is 15pt

\begin{document}

\begin{titlepage}
  \begin{center}
    \Huge
    \textcolor{myred}{Gregorio} and \textcolor{myred}{Gregorio\TeX}:

    Tools for gregorian score engraving.

    \vspace{1cm}

    \large Version \textbf{4.0.0-rc1}, 8 October 2015 %% PARSE_VERSION_DATE

    \vspace{1.5cm}
  \end{center}
  \gresetlinecolor{gregoriocolor}
  \grechangestyle{lowchoralsign}{\cminfamily\small}%
  \grechangestyle{highchoralsign}{\cminfamily\small}%
  \def\GreStar{\greheightstar}%
  \greannotation{\scriptsize{Comm.}}%
  \greannotation{\scriptsize{VII}}%
  \begingroup%
    \color{black!60}%
    \setmainfont[SmallCapsFont=AlegreyaSC]{Alegreya}%
    \newlength{\mini}%
    \setlength{\mini}{\hsize}%
    \addtolength{\mini}{-4cm}%
    \setlength{\fboxsep}{5mm}%
    \hfill\fbox{\parbox{\mini}{\gregorioscore[f]{factus}}}\hfill%
  \endgroup
  \begin{center}
    \vspace{1.5cm}%
    \href{http://gregorio-project.github.io/}{Homepage}

    Source code available on
    \href{http://github.com/gregorio-project/gregorio}{GitHub}.
  \end{center}

  \vspace{2cm}
\end{titlepage}

\cleardoublepage

\tableofcontents

\setlength{\parskip}{\bigskipamount}
\cleardoublepage

\section{Gregorio\TeX{} Macros}
The following sections document the macros available in the Gregorio\TeX{} package. The format is as follows:

\macroname{MacroName}{\{Args\}}{Source File}
Description of macro.

\begin{argtable}
  Arg \# & Data type & Description of argument\\
\end{argtable}

The source file where the macro is defined is included for developers
who wish to consult it.

Some of the macros intended for inclusion in the main.tex file by the user
include usage examples.

Macros are divided into three groups:
\begin{description}
\item[User Commands] These macros are meant to be used by your average user in their \TeX\ files in order to fine tune the appearance of their scores.  They should have names which consist solely of lowercase letters and be prefixed with the \texttt{gre}.  Where the name clearly identifies the function as belonging to Gregorio\TeX, the prefix may be omitted.
\item[Gregorio Controls] These macros are written by the command line tool to gtex files and should not appear outside of gtex files.  They should have names which are in CamelCase and be prefixed with \texttt{Gre}.
\item[Gregorio\TeX\ internals] These macros are used by Gregorio\TeX\ to process and typeset a score and should not appear in a user's document anywhere (not even in gtex files).  They should have names which are all lowercase and be prefixed with \makeatletter\texttt{gre@}\makeatother.
\end{description}


\input{Command_Index_User}
\input{Command_Index_gregorio}
\input{Command_Index_internal}
\input{Gabc}

\begin{appendices}
  \renewcommand\thetable{\thesection\arabic{table}}
  \renewcommand\thefigure{\thesection\arabic{figure}}
  \input{Appendix_Font_Tables}
\end{appendices}

\addcontentsline{toc}{section}{Index}
\printindex

\end{document}
